---
title: "Thesis data analysis"
format: html
editor: visual
---

```{r}
#| label: load in libraries and data 
#| echo: false

library(tidyverse)
library(osfr) # for loading data in
library(purrr) #for parsing raw df
library(jsonlite) # for parsing raw df
library(tibble) # for parsing
library(stringr) # also for parsing 

# project is public, so don't need authentication

# retrieve the project node and download CSVs
project_node <- osf_retrieve_node("w6rxq")

# create a local directory for the data
local_dir <- "data/osf_csvs"
if (!dir.exists(local_dir)) dir.create(local_dir, recursive = TRUE)

# list all files (n_max = Inf ensures we get all ~100 files)
# and download them to the local directory
osf_files <- project_node %>%
  osf_ls_files(n_max = Inf) %>%
  filter(str_detect(name, "\\.csv$"))

osf_download(osf_files, path = local_dir, conflicts = "skip")

# read and combine all CSVs into one dataframe
csv_paths <- list.files(local_dir, pattern = "\\.csv$", full.names = TRUE)

raw_df <- csv_paths %>%
  set_names(basename(.) %>% str_remove("\\.csv$")) %>%
  map_dfr(~read_csv(.x, show_col_types = FALSE), .id = "csv_name")

```

```{r}
#| label: turn response data into question and answer columns
#| echo: false

# keeping all rows but parsing the response column to separate each question into its own row with an answer column
parsed_long_df <- raw_df %>%
  mutate(
    parsed_response = map(response, ~ {
      if (!is.na(.x) && str_starts(.x, "\\{")) {
        tryCatch(fromJSON(.x), error = function(e) NULL)
      } else {
        NULL
      }
    })
  ) %>%
  
  mutate(
    parsed_response = map(parsed_response, ~ {
      if (is.null(.x)) {
        tibble(name = NA_character_, value = NA_character_)
      } else {
        enframe(.x) %>%
          mutate(value = as.character(value))
      }
    })
  ) %>%
  
  unnest(parsed_response) %>%
  
  rename(
    question = name,
    answer   = value
  ) 

#check_working <- parsed_long_df %>% select(response, question, answer)
```

```{r}
#| label: clean and separate df into surveys and game 
#| echo: false 

#add condition (ai blind vs aware) column before splitting up 
parsed_long_df <- parsed_long_df %>%
  group_by(prolific_subject_id) %>%
  mutate(condition = ifelse(any(phase == "ai-aware-question"),
                            "aware",
                            "blind")) %>%
  ungroup() %>%
  mutate(prolific_subject_id = ifelse(
   csv_name == "1y0okya8tt", "6926452ca418f7b008a000a3", ifelse(
     csv_name == "uqe14unhtx", "697a9cc58e44ed0a6d497a73", prolific_subject_id))
 ) # manually impute the very first two participants' prolific subject ids (matched by completion time)

## survey df
# create survey df 
surveys_df <- parsed_long_df %>%
  select(csv_name, prolific_subject_id, question, answer, phase, condition, trial_index, trial_type, stimulus, rt, time_elapsed, internal_node_id) %>% # removing cols about game
  filter(question != "question1") %>% # this was text that showed participants the game concept, but it was inside a survey so it was recorded as though it was a survey question 
  filter(!(is.na(question))) # don't need rows that aren't questions

# create game df
game_df <- parsed_long_df %>%
  filter(trial_type == "plugin-game-iframe") %>% # filters only for rows about the actual game 
  mutate(game_order = case_when(  # trial_index tells us whether they played the game first, second, or third!
    trial_index == 3 ~ "Game_1", 
    trial_index == 7 ~ "Game_2",
    trial_index == 11 ~ "Game_3",
    .default = NA
  )) %>%
  mutate(game = phase) %>% # adding this for clarity but keeping phase in the df in case i combine dfs in the future
  select(prolific_subject_id, game, game_data, exit_reason, exit_stage, exit_progress, emergency_text, keystrokes, mouse_events, game_order, condition) 

```

```{r}
#| label: write df csvs just because
#| echo: false 

write.csv(surveys_df, "cleaned_evaluation_survey_data.csv")
write.csv(game_df, "cleaned_game_data.csv")

```

```{r}
#| label: checking frequency of each game 

# make df of just the games
game_frequency_individual_df <- game_df %>%
  group_by(game, condition) %>%
  count()

game_frequency_by_llm_df <- game_df %>%
  mutate(llm = case_when(
    str_detect(game, "gemini") ~ "gemini",
    str_detect(game, "claude") ~ "claude",
    str_detect(game, "chat") ~ "chat",
    .default = NA
  )) %>%
  group_by(llm, condition) %>%
  count()
```

Each game was played a relatively equal number of times individually and by LLM and level (AI-aware vs. AI-blind), with most games evaluated 7 times.

```{r}
#| label: identify which game blocks have been played 
#| echo: false 

##prep dfs 
# turn games df into game blocks format so we can compare with the planned blocks 
game_blocks_played_df <- game_df %>%
  select(prolific_subject_id, game, game_order, condition) %>%
  pivot_wider(names_from = game_order, values_from = game) %>% # turn hyphens into underscores to match entries in the planned blocks
  mutate(across(starts_with("Game_"),
                ~ str_replace_all(., "-", "_")))

# read in the plan for the game blocks 
game_blocks_plan_df <- read.csv("game_blocks.csv") %>%
  rename("block" = "X")

# duplicate the blocks so there's one per condition
game_blocks_plan_w_condition_df <- game_blocks_plan_df[rep(seq_len(nrow(game_blocks_plan_df)), each = 2), ]

#add condition column and update manually changed blocks
game_blocks_plan_w_condition_df <- game_blocks_plan_df %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(condition = rep(c("aware", "blind"), times = nrow(game_blocks_plan_df))) %>%
  mutate( # i changed a game in block 33 before it was played because the game that was previously in this block (game_chat_4_decibel) was going to get overplayed 
    Game_2 = ifelse(block == 33,
                    "game_chat_1_echo",
                    Game_2)
  ) %>%
  mutate( # i changed a game in block 9 before it was played because the game that was previously in this block was going to get overplayed 
    Game_2 = ifelse(block == 9,
                    "game_claude_5_cartographer",
                    Game_2)
  ) %>%
  mutate(Game_3 = ifelse(block == 11 & condition == "aware", "game_claude_1_light", Game_3)) %>%
  mutate(Game_2 = ifelse(block == 7 & condition == "aware", "game_claude_1_light", Game_2)) %>%
  mutate(Game_1 = ifelse(block == 7 & condition == "blind", "game_chat_1_echo", Game_1)) %>%
  mutate(Game_1 = ifelse(block == 17 & condition == "aware", "game_chat_5_glider", Game_1)) %>%
  mutate(Game_1 = ifelse(block == 18 & condition == "blind", "game_chat_1_echo", Game_1)) %>%
  mutate(Game_2 = ifelse(block == 18 & condition == "blind", "game_claude_2_moth", Game_2)) %>%
  mutate(Game_3 = ifelse(block == 20 & condition == "blind", "game_claude_1_light", Game_3)) 
### add in manipulations of blocks 28 blind, 37 blind, 40 blind here


## now actually check how many times each block has been played
# join the dfs to label the blocks that participants actually played 
game_blocks_played_labeled_df <- game_blocks_played_df %>%
  left_join(
    game_blocks_plan_w_condition_df,
    by = c("Game_1", "Game_2", "Game_3", "condition")
  )

# now count how many times a block was played per condition
block_counts <- game_blocks_played_labeled_df %>%
  count(block, condition, name = "times_played") %>%
  arrange(block)

# merge w full df to see which haven't been played at all
block_counts_full <- game_blocks_plan_w_condition_df %>%
  distinct(block, condition) %>%
  left_join(block_counts, by = c("block", "condition")) %>%
  mutate(times_played = replace_na(times_played, 0)) %>%
  arrange(block)

blocks_to_remove_from_timeline <- block_counts_full %>%
  filter(times_played != 0)

blocks_not_played <- block_counts_full %>%
  filter(times_played == 0)
blocks_not_played

```


```{r}
#| label: look at emergency exits 
#| echo: false 

emergency_exits <- game_df %>%
  filter(exit_reason != "game-complete") %>%
  select(prolific_subject_id, game, emergency_text, game_order)

#grab count. ideally want it to be 0
n_emergency_exits <- emergency_exits %>%
  nrow()

```

Participants exited a game early `r n_emergency_exits` times, presumably due to technical errors. 

```{r}
#| label: attention and commitment check
#| echo: false 

## attention check 
# make df of people who failed 
attention_check_flagged <- surveys_df %>%
  filter(question == "attention-check") %>%
  group_by(prolific_subject_id) %>%
  mutate(n_pass = sum(answer == 3),
    flag = case_when(
      n_pass == 0 ~ "fail (triple)", # don't pay on prolific
      n_pass == 1 ~ "fail (double)", # don't pay on prolific
      n_pass == 2 ~ "fail (single)", # do pay on prolific
      n_pass == 3 ~ "pass", # yay 
      TRUE ~ NA_character_ )) %>%
  ungroup() %>%
  filter(flag != "pass")

#extract list of ids so we can remove their data
failed_attention_check_ids <- attention_check_flagged %>%
  distinct(prolific_subject_id) %>%
  pull(prolific_subject_id)

## commitment check
# make df 
fail_commitment_check <- surveys_df %>%
  filter(question == "commitment-check") %>%
  filter(answer != "yes")

#extract list of ids so we can remove their data
failed_commitment_check_ids <- fail_commitment_check %>%
  distinct(prolific_subject_id) %>%
  pull(prolific_subject_id)


# now remove from df 
surveys_df <- surveys_df %>%
  filter(!prolific_subject_id %in% failed_attention_check_ids) %>%
  filter(!prolific_subject_id %in% failed_commitment_check_ids)
```

```{r}
#| label: count nums of subjects (before cleaning, failed checks, and after cleaning)
#| echo: false

total_num_not_clean_subjects_in_data <- n_distinct(parsed_long_df$prolific_subject_id) 

n_failed_commitment_check <- fail_commitment_check %>%
  summarize(n_failed_commitment_check = n_distinct(prolific_subject_id)) %>%
  pull(n_failed_commitment_check)

n_failed_attention_checks <- attention_check_flagged %>%
  summarize(n_failed_attention_checks = n_distinct(prolific_subject_id)) %>%
  pull(n_failed_attention_checks)

total_num_clean_subjects_in_data <- n_distinct(surveys_df$prolific_subject_id)

```

Through Prolific, 83 participants completed the study. We received data from `r total_num_not_clean_subjects_in_data` of these participants. We excluded data from `r n_failed_attention_checks` subjects for failing one or more of our attention check questions and `r n_failed_commitment_check` subjects for not committing to providing thoughtful answers to the questions in the survey. This resulted in a total of `r total_num_clean_subjects_in_data` participants' data being analyzed.


