---
title: "Thesis data analysis"
format: html
editor: visual
---

This document is the data analysis source code for my pre-registered study, Are Large Language Models Creative? 

Pre-registration can be found at https://osf.io/hw2en/overview. 

This code does not retrieve participant data directly from OSF. The data must be downloaded from the Github repository in which this analysis file is found. They are in the folder osfstorage-archive. 

Coded qualitative data files also must be downloaded from the Github repository in which this analysis file is found. These files are in the folder coded_qual_data. I coded the qualitative data for themes in Google Sheets and then analyzed the overarching trends in this document. 


# Data wrangling & basic sample stats
```{r}
#| label: load in libraries and data 
#| include: false
#| warning: false
#| message: false

library(tidyverse) # for everything
library(jsonlite) # for parsing raw df
library(lme4) # for LMM
library(lmerTest) # for ICC 
library(emmeans) # for marginal means
library(likert)  # for making likert plots
library(knitr) # for making nicer tables 
library(scales) # for using percent function
library(corrplot) # for making correlation matrix 
library(ggpubr) # for plots 
library(factoextra) # for plots 
library(kableExtra) #for table

### ditching direct osf download becuase it downloads a different set of files every time and i can't figure out why

# # retrieve the project node and download CSVs
# project_node <- osf_retrieve_node("w6rxq")
# 
# # create a local directory for the data
# local_dir <- "data/osf_csvs"
# if (!dir.exists(local_dir)) dir.create(local_dir, recursive = TRUE)
# 
# # UNCOMMENT THESE LINES FOR REPRODUCIBILITY
# # list all files (n_max = Inf ensures we get all ~100 files)
# # and download them to the local directory
# osf_files <- project_node %>%
#   osf_ls_files(n_max = Inf) %>%
#   filter(str_detect(name, "\\.csv$"))
# 
# osf_download(osf_files, path = local_dir, conflicts = "skip")
# 
# # read and combine all CSVs into one dataframe
# csv_paths <- list.files(local_dir, pattern = "\\.csv$", full.names = TRUE) %>%
#   sort()
# 
# raw_df <- csv_paths %>%
#   set_names(basename(.) %>% str_remove("\\.csv$")) %>%
#   purrr::map_dfr(readr::read_csv, show_col_types = FALSE, .id = "csv_name")


### manually downloaded osf files folder from the website and added it to this project  
# folder containing all OSF csv files
data_dir <- "osfstorage-archive"

# get all csv paths (sorted for reproducibility)
csv_paths <- list.files(
  data_dir,
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  sort()

# read and combine
raw_df <- csv_paths %>%
  set_names(basename(.) %>% str_remove("\\.csv$")) %>% 
  purrr::map_dfr(readr::read_csv, show_col_types = FALSE, .id = "csv_name")

```


```{r}
#| label: turn response data into question and answer columns
#| include: false
#| warning: false
#| message: false

# keeping all rows but parsing the response column to separate each question into its own row with an answer column
parsed_long_df <- raw_df %>%
  mutate(
    parsed_response = map(response, ~ {
      if (!is.na(.x) && str_starts(.x, "\\{")) {
        tryCatch(fromJSON(.x), error = function(e) NULL)
      } else {
        NULL
      }
    })
  ) %>%
  
  mutate(
    parsed_response = map(parsed_response, ~ {
      if (is.null(.x)) {
        tibble(name = NA_character_, value = NA_character_)
      } else {
        enframe(.x) %>%
          mutate(value = as.character(value))
      }
    })
  ) %>%
  
  unnest(parsed_response) %>%
  
  rename(
    question = name,
    answer   = value
  ) 

#check_working <- parsed_long_df %>% select(response, question, answer)
```

```{r}
#| label: add prolific demo file & merge with df
#| echo: false
#| warning: false
#| message: false

#add condition (ai blind vs aware) column and manually impute some missing ids
parsed_long_df <- parsed_long_df %>%
  group_by(prolific_subject_id) %>%
  mutate(condition = ifelse(any(phase == "ai-aware-question"),
                            "aware",
                            "blind")) %>%
  ungroup() %>%
  mutate(prolific_subject_id = ifelse(
   csv_name == "1y0okya8tt", "6926452ca418f7b008a000a3", ifelse(
     csv_name == "uqe14unhtx", "697a9cc58e44ed0a6d497a73", prolific_subject_id))
 )

demo_df <- read_csv("prolific_demographic_export_6987c292780adf01bb926159.csv") %>%
  rename(prolific_subject_id = "Participant id") # match name to our df

parsed_long_df <- right_join(parsed_long_df, demo_df, by = "prolific_subject_id") %>% # only keeping obvs in both
 filter(!(is.na(csv_name))) %>%
 filter(Status == "APPROVED" | prolific_subject_id == "670ff2e18c2f93beac6a6a72")


testing <- parsed_long_df %>%
  filter(condition == "blind") 

#n_distinct(parsed_long_df$prolific_subject_id)
```

```{r}
#| label: separate df into surveys and game dfs
#| echo: false 
## survey df
# create survey df 

surveys_df <- parsed_long_df %>%
  select(csv_name, prolific_subject_id, question, answer, phase, condition, trial_index, trial_type, stimulus, rt, time_elapsed, internal_node_id, "Highest education level completed", "Age", "Gender", "Video games", "Ethnicity simplified", "Country of residence", "Employment status", "Language", "Highest education level completed", "Student status") %>% # removing cols about game
  filter(question != "question1") %>% # this was text that showed participants the game concept, but it was inside a survey so it was recorded as though it was a survey question 
  filter(!(is.na(question))) # don't need rows that aren't questions

# create game df
game_df <- parsed_long_df %>%
  filter(trial_type == "plugin-game-iframe") %>% # filters only for rows about the actual game 
  mutate(game_order = case_when(  # trial_index tells us whether they played the game first, second, or third!
    trial_index == 3 ~ "Game_1", 
    trial_index == 7 ~ "Game_2",
    trial_index == 11 ~ "Game_3",
    .default = NA
  )) %>%
  mutate(game = phase) %>% # adding this for clarity but keeping phase in the df in case i combine dfs in the future
  select(prolific_subject_id, game, game_data, exit_reason, exit_stage, exit_progress, emergency_text, keystrokes, mouse_events, game_order, condition)

```

```{r}
#| label: attention and commitment check
#| echo: false 

## attention check 
# make df of people who failed 
attention_check_flagged <- surveys_df %>%
  filter(question == "attention-check") %>%
  group_by(prolific_subject_id) %>%
  mutate(n_pass = sum(answer == 3),
    flag = case_when(
      n_pass == 0 ~ "fail (triple)", # don't pay on prolific
      n_pass == 1 ~ "fail (double)", # don't pay on prolific
      n_pass == 2 ~ "fail (single)", # do pay on prolific
      n_pass == 3 ~ "pass", # yay 
      TRUE ~ NA_character_ )) %>%
  ungroup() %>%
  filter(flag != "pass")

#extract list of ids so we can remove their data
failed_attention_check_ids <- attention_check_flagged %>%
  distinct(prolific_subject_id) %>%
  pull(prolific_subject_id)

## commitment check
# make df 
fail_commitment_check <- surveys_df %>%
  filter(question == "commitment-check") %>%
  filter(answer != "yes")

#extract list of ids so we can remove their data
failed_commitment_check_ids <- fail_commitment_check %>%
  distinct(prolific_subject_id) %>%
  pull(prolific_subject_id)

# now remove from df. also remove subject who returned their study after completion. we still have their evaluation data on osf so we could use it; however, their messaging on prolific is stressing me out 
surveys_df <- surveys_df %>%
  filter(!prolific_subject_id %in% failed_attention_check_ids) %>%
  filter(!prolific_subject_id %in% failed_commitment_check_ids) 

game_df <- game_df %>%
  filter(!prolific_subject_id %in% failed_attention_check_ids) %>%
  filter(!prolific_subject_id %in% failed_commitment_check_ids) 

```

```{r}
#| label: count nums of subjects (before cleaning, failed checks, and after cleaning)
#| echo: false

total_num_not_clean_subjects_in_data <- n_distinct(parsed_long_df$prolific_subject_id) 

n_failed_commitment_check <- fail_commitment_check %>%
  summarize(n_failed_commitment_check = n_distinct(prolific_subject_id)) %>%
  pull(n_failed_commitment_check)

n_failed_attention_checks <- attention_check_flagged %>%
  summarize(n_failed_attention_checks = n_distinct(prolific_subject_id)) %>%
  pull(n_failed_attention_checks)

total_num_clean_subjects_in_data <- n_distinct(surveys_df$prolific_subject_id)

```

```{r}
#| label: demographic summary stats
#| echo: false

# creating demo df from surveys_df because i want to include blind column
demo_df <- surveys_df %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  filter(Age != "CONSENT_REVOKED") %>% 
  mutate(Age = as.numeric(Age))

# age
age_mean <- round(mean(demo_df$Age, na.rm = TRUE), 0)
age_sd   <- round(sd(demo_df$Age, na.rm = TRUE), 0)

# gender
gender_dist <- demo_df %>%
  count(Gender) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# video games (categorical)
vg_dist <- demo_df %>%
  count(`Video games`) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# country of residence
country_dist <- demo_df %>%
  count(`Country of residence`, condition, sort = TRUE) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# education
education_dist <- demo_df %>%
  count(`Highest education level completed`, sort = TRUE) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# ethnicity simplified
ethnicity_dist <- demo_df %>%
  count(`Ethnicity simplified`, sort = TRUE) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# first language  ---- i don't think this is really relevant 
language_dist <- demo_df %>%
  count(`Language`, sort = TRUE) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# employment status  --- i don't think this is really relevant 
employment_dist <- demo_df %>%
  count(`Employment status`, sort = TRUE) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

# student status ---- i don't think this is really relevant here
student_dist <- demo_df %>%
  count(`Student status`, sort = TRUE) %>%
  mutate(p = percent(n / sum(n), accuracy = 0.1))

```

```{r}
#| label: write df csvs 
#| echo: false

write.csv(surveys_df, "cleaned_evaluation_survey_data.csv")
write.csv(game_df, "cleaned_game_data.csv")

```

```{r}
#| label: (for conducting experiment) checking frequency of each game 
#| echo: false

# make df of just the games
game_frequency_individual_df <- game_df %>%
  group_by(game, condition) %>%
  count()

game_frequency_by_llm_df <- game_df %>%
  mutate(llm = case_when(
    str_detect(game, "gemini") ~ "gemini",
    str_detect(game, "claude") ~ "claude",
    str_detect(game, "chat") ~ "chat",
    .default = NA
  )) %>%
  group_by(llm, condition) %>%
  count()

#game_frequency_individual_df
#game_frequency_by_llm_df

```

```{r}
#| label: (for conducting experiment) identify which game blocks have been played 
#| include: false 

##prep dfs 
# turn games df into game blocks format so we can compare with the planned blocks 
game_blocks_played_df <- game_df %>%
  select(prolific_subject_id, game, game_order, condition) %>%
  pivot_wider(names_from = game_order, values_from = game) %>% # turn hyphens into underscores to match entries in the planned blocks
  mutate(across(starts_with("Game_"),
                ~ str_replace_all(., "-", "_")))

# read in the plan for the game blocks 
game_blocks_plan_df <- read.csv("game_blocks.csv") %>%
  rename("block" = "X")

# duplicate the blocks so there's one per condition
game_blocks_plan_w_condition_df <- game_blocks_plan_df[rep(seq_len(nrow(game_blocks_plan_df)), each = 2), ]

#add condition column and update manually changed blocks
game_blocks_plan_w_condition_df <- game_blocks_plan_df %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(condition = rep(c("aware", "blind"), times = nrow(game_blocks_plan_df))) %>%
  mutate( # i changed a game in block 33 before it was played because the game that was previously in this block (game_chat_4_decibel) was going to get overplayed 
    Game_2 = ifelse(block == 33,
                    "game_chat_1_echo",
                    Game_2)
  ) %>%
  mutate( # i changed a game in block 9 before it was played because the game that was previously in this block was going to get overplayed 
    Game_2 = ifelse(block == 9,
                    "game_claude_5_cartographer",
                    Game_2)
  ) %>%
  mutate(Game_3 = ifelse(block == 11 & condition == "aware", "game_claude_1_light", Game_3)) %>%
  mutate(Game_2 = ifelse(block == 7 & condition == "aware", "game_claude_1_light", Game_2)) %>%
  mutate(Game_1 = ifelse(block == 7 & condition == "blind", "game_chat_1_echo", Game_1)) %>%
  mutate(Game_1 = ifelse(block == 17 & condition == "aware", "game_chat_5_glider", Game_1)) %>%
  mutate(Game_1 = ifelse(block == 18 & condition == "blind", "game_chat_1_echo", Game_1)) %>%
  mutate(Game_2 = ifelse(block == 18 & condition == "blind", "game_claude_2_moth", Game_2)) %>%
  mutate(Game_3 = ifelse(block == 20 & condition == "blind", "game_claude_1_light", Game_3)) 
### add in manipulations of blocks 28 blind, 37 blind, 40 blind here


## now actually check how many times each block has been played
# join the dfs to label the blocks that participants actually played 
game_blocks_played_labeled_df <- game_blocks_played_df %>%
  left_join(
    game_blocks_plan_w_condition_df,
    by = c("Game_1", "Game_2", "Game_3", "condition")
  )

# now count how many times a block was played per condition
block_counts <- game_blocks_played_labeled_df %>%
  count(block, condition, name = "times_played") %>%
  arrange(block)

# merge w full df to see which haven't been played at all
block_counts_full <- game_blocks_plan_w_condition_df %>%
  distinct(block, condition) %>%
  left_join(block_counts, by = c("block", "condition")) %>%
  mutate(times_played = replace_na(times_played, 0)) %>%
  arrange(block)

blocks_to_remove_from_timeline <- block_counts_full %>%
  filter(times_played != 0)

blocks_not_played <- block_counts_full %>%
  filter(times_played == 0)
blocks_not_played

```

# Emergency exits

```{r}
#| label: look at emergency exits (quantitative)
#| include: false 
#| warning: false
#| message: false

#make df 
emergency_exits <- game_df %>%
  filter(exit_reason != "game-complete") %>%
  select(prolific_subject_id, game, emergency_text, game_order)

#grab count
n_emergency_exits <- emergency_exits %>%
  nrow()

# see which games were exited and how many times 
n_emergency_exits_per_game <- emergency_exits %>%
  count(game) %>%
  arrange(desc(n)) 

# see how often a participant exited 0 games, just 1 game, 2 games, or all 3
n_emergency_exits_per_participant <- emergency_exits %>%
  count(prolific_subject_id) %>%
  #arrange(desc(n)) # can cut off code here if i want to see who exited 
  rename(n_games_exited_by_participant = n) %>%
  count(n_games_exited_by_participant) %>%
  rename(n_participants = n)
  
kable(n_emergency_exits_per_game, col.names = c("Game", "Number of exits from game"), format = "html")


# see whether game order seemed to be at play
n_emergency_exits_by_game_order <- emergency_exits %>%
  count(game_order) %>%
  #arrange(desc(n)) # can cut off code here if i want to see who exited 
  rename(n_exits = n) %>%
  count(n_exits) %>%
  rename(n_participants = n)
  
```

```{r}
#| label: emergency exits qual analysis
#| include: false 

#write csv so i can do encoded analysis in excel
write.csv(emergency_exits, "emergency_exits.csv")

#load analysis back in. in this analysis, i tracked peoples' reported reasons for exiting the game
emergency_exit_qual_raw <- read_csv("emergency_exits_qual_encoded.csv")

# pull exit progress to put back in the qual df since i didn't include this when i first wrote the emergency exits df that i then added qual analysis to. i didn't put a lot of effort into making the exit progress tracking, so it might turn out to be pretty unhelpful
exit_progress_for_merging <- game_df %>%
  select(prolific_subject_id, condition, game, exit_stage, exit_progress, keystrokes, mouse_events)

emergency_exit_qual <- emergency_exit_qual_raw %>%
  merge(exit_progress_for_merging, by = c("prolific_subject_id", "game")) %>%
  select(prolific_subject_id, condition, game, ambiguity_level, genuine_fatal_game_bug, alleged_fatal_game_bug, too_difficult, gameplay_unclear, no_reason_apparent_lack_of_problems, stuck_unclear, insufficient_information, game_order, emergency_text, exit_stage, exit_progress, keystrokes, mouse_events, my_summary) 

#now looking at the stage in the game at which the participants exited. the data for this isn't as clear so i can't quite tell with some people.

# the exit progress for some of them would have me think that the person exited at the start of the game but then their exit reason cites things like not being able to get through a part of the game that isn't at the start, so i'm a little confused. i think some of the exit progress trackers aren't trustworthy, so i maybe shouldn't pursue this. it wouldn't be super informative anyway; would just be nice to know 


### for inline text: i'm just manually writing it in the report for ease and since i'm not using this actual doc in my thesis

#by reason:
genuine_bug_total_exits <- emergency_exit_qual %>% filter(genuine_fatal_game_bug == 1) %>% count
genuine_bug_total_games <- emergency_exit_qual %>% filter(genuine_fatal_game_bug == 1) %>% summarise(n = n_distinct(game))

alleged_bug_total_exits <- emergency_exit_qual %>% filter(alleged_fatal_game_bug == 1) %>% count
alleged_bug_total_games <- emergency_exit_qual %>% filter(alleged_fatal_game_bug == 1) %>% summarise(n = n_distinct(game))

difficulty_total_exits <- emergency_exit_qual %>% filter(too_difficult == 1) %>% count
difficulty_total_games <- emergency_exit_qual %>% filter(too_difficult == 1) %>% summarize(n = n_distinct(game))

gameplay_unclear_total_exits <- emergency_exit_qual %>% filter(gameplay_unclear == 1) %>% count
gameplay_unclear_total_games <- emergency_exit_qual %>% filter(gameplay_unclear == 1) %>% summarise(n = n_distinct(game))

no_problem_total_exits <- emergency_exit_qual %>% filter(no_reason_apparent_lack_of_problems == 1) %>% count
no_problem_total_games <- emergency_exit_qual %>% filter(no_reason_apparent_lack_of_problems == 1) %>% summarize(n = n_distinct(game))

insufficient_info_total_exits <- emergency_exit_qual %>% filter(insufficient_information == 1 | stuck_unclear == 1) %>% count

insufficient_info_total_games <- emergency_exit_qual %>% filter(insufficient_information == 1 | stuck_unclear == 1) %>% summarize(n = n_distinct(game))

emergency_exit_qual %>% filter(ambiguity_level == "High") %>% group_by(game) %>% count

genuine_bug_total_exits # good
genuine_bug_total_games # good 
alleged_bug_total_exits
alleged_bug_total_games
difficulty_total_exits # good 
difficulty_total_games # good 
gameplay_unclear_total_exits # good
gameplay_unclear_total_games # good
no_problem_total_exits # good 
no_problem_total_games # good 
insufficient_info_total_exits # good
insufficient_info_total_games # good 

```

```{r}
#| label: old exit summary text
#| echo: false 
# Across the `r total_num_clean_subjects_in_data*3` total gameplays (`r total_num_clean_subjects_in_data` subjects x 3 gameplays per subject), there were 41 emergency exits involving 14 of the 18 video games. `r genuine_bug_total_exits + alleged_bug_total_exits + difficulty_total_exits + gameplay_unclear_total_exits` of these exits cited playability issues.
# 
# No participant exited all three of their assigned games, and the large majority of exits (`r genuine_bug_total_exits + alleged_bug_total_exits + difficulty_total_exits + gameplay_unclear_total_exits` of 41) cited playability issues of bugs, unclear gameplay, or extreme difficulty. Among the remaining exits, 6 indicated an inability to complete or exit the game but did not specify a reason (e.g., bug, difficulty, or unclear gameplay), making it impossible to determine whether these reflected genuine technical problems or other concerns. The other two exits seemed to indicate that the participant had successfully completed the game without error. Taken together, the exit data do not suggest systematic misuse of the emergency exit button.

```


```{r}
#| label: old emergency exit qual result report text
#| echo: false
#| eval: false


# One of the top-exited games (Claude-1) was exited six times, half by AI-blind and half by AI-aware participants. Four exits cited the game being too difficult, one exit indicated an inability to complete and exit the game but did not explain why (e.g., bug, too difficult, gameplay unclear), and one (AI-aware) exit cited an alleged bug that, based on their description, was likely one of the intentional gameplay features.
# 
# The other equally top-exited game (Gemini-4) was exited mostly by AI-blind participants. Half (three) of the exits were due to unclear gameplay, two cited the game being too difficult, and one (AI-blind) suggested there may have been a bug because they "couldn't do anything".
# 
# Another game (Chat 2) was exited four times, half by AI-blind and half by AI-aware participants. Three cited game bugs. Two of these citations were for the same problem; I and the other participants who completed this game did not encounter this issue, but nonetheless it likely exits given that both subjects encountered this issue. The third bug citation was for a bug that indeed exists but does not prevent the game from being playable. The fourth exit indicated an inability to complete and exit the game but did not explain why (e.g., bug, too difficult, gameplay unclear).
# 
# Another game (Chat-3) was exited three times, mostly by aware participants. One cited high difficulty, one seemed to have completed the game properly with no issues, and one indicated an inability to complete and exit the game but did not explain why (e.g., bug, too difficult, gameplay unclear).
# 
# Another game (Chat-6) was exited three times, mostly by blind participants. One cited high difficulty, one cited unclear gameplay, and one cited an alleged bug that may have been an actual bug or may have been a misunderstanding of the game (I'm genuinely not sure).
# 
# Another game (Claude-5) was exited three times, all by AI-aware participants. One cited unclear gameplay, one indicated an inability to complete and exit the game but did not explain why (e.g., bug, too difficult, gameplay unclear), and one cited an alleged bug that almost certainly was actually caused by the participant not reading the instruction to click "Enter" after completing the game so that they can proceed with the study rather than playing the game again.
# 
# Two of the games (Gemini-5 & Claude-6), each had only one exit. The participants reported a gameplay issue (that, presumably per our instructions, they felt prevented them from completing the game); however, the issues that they faced were intentional and correctly functioning challenges in the game, suggesting the actual problem is that the game play was unclear to participants. Both participants were AI-aware, which is interesting---they may have been more quick to perceive the challenges in the games (especially if they were unclear) as bugs because they knew the games weren't designed by humans. This event is extra interesting for the Claude game since the issue that the participant faced was designed to be confusing in the moment but then make sense in an emotionally resonant way at the end of the game; so the participant didn't wait to 'hear out' the game. Also, if they had waited/paid attention for even just ten seconds, they would have noticed that the 'issue' was actually still progressing the game to completion, so it really seems like they didn't wait to see if things would pan out.
# 
# One of the games (Gemini-6) was exited once in total due to the participant finding the gameplay unclear.
# 
# One game (Gemini-2) was exited once in total due to the participant encountering some type of issue in completing and exiting the game, but they did not specify a reason (e.g., bug, too difficult, gameplay unclear).
# 
# Two games (Chat-4, Claude-2) were each exited once but the exit reasons provided insufficient information.
```

# More data wrangling & basic quantitative stats
```{r}
#| label: make dfs for qual analysis 
#| echo: false 

surveys_qualitative_df <- surveys_df %>%
  filter(str_detect(question, "free-response|ai_blind_question|ai_aware_question")) %>%
  select(prolific_subject_id, condition, phase, question, answer)

creativity_qual_df <- surveys_qualitative_df %>%
  filter(question == "creativity-free-response") %>%
  select(prolific_subject_id, question, answer)

write.csv(creativity_qual_df, "creativity_qual_data.csv")

originality_qual_df <- surveys_qualitative_df %>%
  filter(question == "originality-free-response") %>%
  select(prolific_subject_id, question, answer)

write.csv(originality_qual_df, "originality_qual_data.csv")

effectiveness_qual_df <- surveys_qualitative_df %>%
  filter(question == "effectiveness-free-response") %>%
  select(prolific_subject_id, question, answer)

write.csv(effectiveness_qual_df, "effectiveness_qual_data.csv")

intentionality_qual_df <- surveys_qualitative_df %>%
  filter(question == "intentionality-free-response") %>%
  select(prolific_subject_id, question, answer)

write.csv(intentionality_qual_df, "intentionality_qual_data.csv")

authenticity_intentionality_qual_df <- surveys_qualitative_df %>%
  filter(question == "authenticity-intentionality-free-response") %>%
  select(prolific_subject_id, question, answer)

write.csv(authenticity_intentionality_qual_df, "authenticity_intentionality_qual_data.csv")

ai_blind_qual_df <- surveys_qualitative_df %>%
  filter(question == "ai_blind_question") %>%
  select(prolific_subject_id, question, answer)

write.csv(ai_blind_qual_df, "ai_blind_data.csv")

ai_aware_qual_df <- surveys_qualitative_df %>%
  filter(question == "ai_aware_question") %>%
  select(prolific_subject_id, question, answer)

write.csv(ai_aware_qual_df, "ai_aware_qual_data.csv")

```

```{r}
#| label: make various dfs for quant analyses
#| echo: false

# convert answer to numeric for rating questions 
surveys_numeric_df <- surveys_df %>%
  filter(str_detect(question, "rating")) %>%
  mutate(answer = as.numeric(answer))

# extract baseline ratings (one per participant)
baseline_df <- surveys_numeric_df %>%
  filter(phase == "baseline-questions-rating") %>%
  select(prolific_subject_id, question, answer) %>%
  pivot_wider(names_from = question, values_from = answer)

# extract game evaluation ratings 
# (phase looks like "evaluation-chat-1-echo")
evaluation_df <- surveys_numeric_df %>%
  filter(str_detect(phase, "^evaluation-")) %>%
  mutate(game_id = str_remove(phase, "^evaluation-")) %>%
  select(prolific_subject_id, game_id, condition, question, answer) %>%
  pivot_wider(names_from = question, values_from = answer)

# extract concept evaluation ratings
# (phase looks like "concept-chat-1-echo")
concept_df <- surveys_numeric_df %>%
  filter(str_detect(phase, "^concept-")) %>%
  mutate(game_id = str_remove(phase, "^concept-")) %>%
  select(prolific_subject_id, game_id, question, answer) %>%
  pivot_wider(names_from = question, values_from = answer)

# join everything together into one analysis dataframe
clean_analysis_df <- evaluation_df %>%
  left_join(concept_df, by = c("prolific_subject_id", "game_id")) %>%
  left_join(baseline_df, by = "prolific_subject_id") %>%
  # Clean up column names for easier modeling
  rename(
    creativity = `creativity-rating`,
    originality = `originality-rating`,
    effectiveness = `effectiveness-rating`,
    authenticity = `authenticity-rating`,
    intentionality = `intentionality-rating`,
    concept_creativity = `concept-creativity-rating`,
    implementation_accuracy = `concept-implementation-rating`
  )

difference_ratings_df <- clean_analysis_df %>%
  mutate(difference_originality = 
           originality - baseline_originality_rating) %>%
  mutate(difference_effectiveness = 
           effectiveness - baseline_effectiveness_rating) %>%
  mutate(difference_authenticity = 
           authenticity - baseline_authenticity_rating) %>%
  mutate(difference_intentionality = 
           intentionality - baseline_intentionality_rating) 

```

```{r}
#| label: basic summary stats on ratings (not sure if these are actually useful)
#| echo: false 
#| message: false
#| warning: false 

ratings_summary_stats <- surveys_numeric_df %>%
  group_by(condition, question) %>%
  summarize(
    n = n(),
    mean_rating = round(mean(answer, na.rm = TRUE), 2),
    sd_rating = round(sd(answer, na.rm = TRUE), 2),
    se_rating = round(sd(answer, na.rm = TRUE) / sqrt(n()), 2),
    median_rating = median(answer, na.rm = TRUE),
    q25 = quantile(answer, 0.25, na.rm = TRUE),
    q75 = quantile(answer, 0.75, na.rm = TRUE),
    min_rating = min(answer, na.rm = TRUE),
    max_rating = max(answer, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = condition,
    values_from = c(mean_rating, sd_rating, median_rating, n)
  )

```

```{r}
#| label: extract percent of participants giving each rating to each item 
#| echo: false

# compute percentage of responses at each level for each question
pct_ratings_per_level_and_item <- surveys_numeric_df %>%
  group_by(question, answer) %>%             # group by question and rating level
  summarize(n = n(), .groups = "drop") %>%   # count how many responses at each level
  group_by(question) %>%
  mutate(pct = n / sum(n) * 100) %>%        # calculate percentage
  ungroup() %>%
  pivot_wider(
    names_from = c(question, answer),        # e.g., creativity-rating_1, creativity-rating_2
    values_from = pct,
    names_sep = "_"
  )

# compute percentage of responses at each level for each question
pct_ratings_per_level_and_item <- surveys_numeric_df %>%
  group_by(question, answer) %>%             # group by question and rating level
  summarize(n = n(), .groups = "drop") %>%   # count how many responses at each level
  group_by(question) %>%
  mutate(pct = n / sum(n) * 100) %>%        # calculate percentage
  ungroup() %>%
  mutate(question = paste0(question, "_", answer))

```


```{r}
#| label: labels for rating items and likert levels
#| echo: false 

rating_item_labels <- c(
  "creativity-rating" = "Creativity",
  "effectiveness-rating" = "Effectiveness",
  "intentionality-rating" = "Intentionality",
  "originality-rating" = "Originality",
  "authenticity-rating" = "Authenticity", 
                   "baseline_authenticity_rating" = "Baseline Authenticity",
                   "baseline_effectiveness_rating" = "Baseline Effectiveness",
                   "baseline_intentionality_rating" = "Baseline Intentionality",
                   "baseline_originality_rating" = "Baseline Originality", 
                   "concept-creativity-rating" = "Concept Creativity", 
                   "concept-implementation-rating" = "Concept Implementation"
                   )

likert_levels <- c("Not at all", "Somewhat", "Moderately", "Very", "Extremely")

```

```{r}
#| label: grab aware and blind sub ids 
#| echo: false

# useful later on for filtering purposes

aware_ids <- surveys_df %>%
  filter(condition == "aware") %>%
  pull(prolific_subject_id)

blind_ids <- surveys_df %>%
  filter(condition == "blind") %>%
  pull(prolific_subject_id)

```

# Inter-judge reliability (ICC)

```{r}
#| label: ICC reliability analysis (per game per creativity dimension)
#| include: false

### THIS ANALYSIS IS SINGULAR. NOT USING. 
## i kept the code just for reference

# moving forward, i did ICC over per game within each condition, but not per creativity dimension. 

#another idea was to have a hierarchical term that is interaction between game and the kind of rating that it is (this moves away from agreement tho)


# function to calculate ICC 
calc_lmm_icc_obsolete <- function(df, outcome) {
  formula <- as.formula(paste(outcome, "~ 1 + (1 | game_id) + (1 | prolific_subject_id)"))
  model <- lmerTest::lmer(formula, data = df)
  vc <- as.data.frame(VarCorr(model))
  
  vg <- vc$vcov[vc$grp == "game_id"]
  vs <- vc$vcov[vc$grp == "prolific_subject_id"]
  vr <- vc$vcov[vc$grp == "Residual"]
  
  # ICC(2,1): how much of a game's rating variance is due to the game itself
  # vs. rater bias + noise — this answers "do raters agree on a given game?"
  icc_single <- vg / (vg + vs + vr)
  
  # ICC(2,k): reliability of the *average* rating across your ~6-7 raters
  # this is more relevant for your use case — how reliable is the mean score
  # you'll actually use in analysis?
  k <- df %>% group_by(game_id) %>% summarise(n = n()) %>% pull(n) %>% mean()
  icc_average <- (k * vg) / (k * vg + vs + vr)
  
  list(icc_single = icc_single, icc_average = icc_average)
}

# ratings to evaluate
ratings_to_check <- c("creativity", "originality", "effectiveness", "authenticity", "intentionality", "concept_creativity", "implementation_accuracy")

# Calculate ICC for each condition
icc_results <- expand_grid(
  condition_val = c("aware", "blind"),
  rating_val = ratings_to_check
) %>%
  mutate(
    icc_out = map2(condition_val, rating_val, ~ {
      df_subset <- clean_analysis_df %>% filter(condition == .x)
      calc_lmm_icc_obsolete(df_subset, .y)
    }),
    icc_single  = map_dbl(icc_out, "icc_single"),
    icc_average = map_dbl(icc_out, "icc_average")
  ) %>%
  select(-icc_out)

```

```{r}
#| label: ICC (per game, all creativity items)
#| echo: false 

# pivot to long format so all rating items are stacked
long_df <- clean_analysis_df %>%
  pivot_longer(
    cols = c(creativity, originality, effectiveness, authenticity, 
             intentionality, concept_creativity, implementation_accuracy),
    names_to  = "rating_item",
    values_to = "rating"
  )

calc_lmm_icc_pooled <- function(df) {
  # rating_item is now a fixed effect (we want item-specific intercepts)
  # rater variance estimated across all items together — more data, stable estimates
  model <- lmerTest::lmer(
    rating ~ rating_item + (1 | prolific_subject_id),
    data = df
  )
  
  vc <- as.data.frame(VarCorr(model))
  vs <- vc$vcov[vc$grp == "prolific_subject_id"]
  vr <- vc$vcov[vc$grp == "Residual"]
  
  k <- df %>% 
    group_by(prolific_subject_id) %>% 
    summarise(n = n()) %>% 
    pull(n) %>% 
    mean()
  
  icc_single  <- vs / (vs + vr)
  icc_average <- vs / (vs + vr/k)
  
  list(
    icc_single      = icc_single,
    icc_average     = icc_average,
    k_avg           = k,
    var_rater       = vs,
    var_residual    = vr
  )
}

# run one model per condition per game (grouping all rating dimensions together)
icc_results <- expand_grid(
  condition_val = c("aware", "blind"),
  game_val      = unique(long_df$game_id)
) %>%
  mutate(
    icc_out = map2(condition_val, game_val, ~ {
      df_subset <- long_df %>% filter(condition == .x, game_id == .y)
      calc_lmm_icc_pooled(df_subset)
    }),
    icc_single   = map_dbl(icc_out, "icc_single"),
    icc_average  = map_dbl(icc_out, "icc_average"),
    k_avg        = map_dbl(icc_out, "k_avg"),
    var_rater    = map_dbl(icc_out, "var_rater"),
    var_residual = map_dbl(icc_out, "var_residual")
  ) %>%
  select(-icc_out)

# summary across games, by condition
icc_summary <- icc_results %>%
  group_by(condition_val) %>%
  summarise(
    mean_icc_single  = mean(icc_single),
    sd_icc_single    = sd(icc_single),
    mean_icc_average = mean(icc_average),
    sd_icc_average   = sd(icc_average),
    .groups = "drop"
  )

poor_icc_single_game <- icc_results %>% 
  filter(icc_single < 0.50)

poor_icc_single_game_summary <- poor_icc_single_game %>%
  summarize(
            min = min(icc_single),
            max = max(icc_single)
            )
```

```{r}
#| label: old icc report text
#| include: false
#| echo: false

# `r count(poor_icc_single_game)` games had poor single-rater ICC ratings (`r round(poor_icc_single_game_summary$min, 2)`-`r round(poor_icc_single_game_summary$max, 2)`), `r nrow(poor_icc_single_game[poor_icc_single_game$condition_val == "blind", ])` of which were from the blind condition. I'm not sure how to interpret this.
# 
# On average, single-rater ICC was acceptable in both the aware condition (ICC(2,1) = `r round(icc_summary$mean_icc_single[icc_summary$condition_val == "aware"], 2)`) and blind condition (ICC(2,1) = `r round(icc_summary$mean_icc_single[icc_summary$condition_val == "blind"], 2)`). 

# Inter-rater reliability was assessed using ICC(2,k) estimated from linear mixed models fit separately for each game within each condition, with rating item as a fixed effect and rater as a random effect. Reliability of the mean rating across the approximately seven raters per game per condition was high in both conditions (aware: ICC(2,k) = `r round(icc_summary$mean_icc_average[icc_summary$condition_val == "aware"], 2)`; blind: ICC(2,k) = `r round(icc_summary$mean_icc_average[icc_summary$condition_val == "blind"], 2)`), supporting the use of mean ratings in subsequent analyses.


```

# Likert creativity and related dimensions
```{r}
#| label: fig-likert-absolute-ratings
#| include: false
#| warning: false
#| message: false
 

# likert plot! 
# put df into likert format 
likert_data_absolute <- surveys_df %>%
  select(phase, prolific_subject_id, question, answer) %>%
  mutate(phase = gsub("concept", "evaluation", phase)) %>%
  filter(str_detect(question, "rating")) %>%
  filter(str_detect(question, "baseline", negate = TRUE)) %>%
  mutate(answer = factor(answer, 
                         levels = 1:5, 
                         labels = likert_levels)) %>%
  pivot_wider(
    id_cols = c(phase, prolific_subject_id),
    names_from = question, 
    values_from = answer) %>% # keep only question columns %>%
  select(-c(phase, prolific_subject_id)) %>%
  mutate(across(everything(),
                ~ factor(.x, levels = likert_levels)))

#plot
likert_plot_all <- plot(likert(as.data.frame(likert_data_absolute)), legend.position="right")
```

```{r}
#| label: fig-likert-absolute-ratings-by-condition
#| include: false
#just for ai aware 
likert_data_absolute_aware <- surveys_df %>%
  filter(condition == "aware") %>%
  select(phase, prolific_subject_id, question, answer) %>%
  mutate(phase = gsub("concept", "evaluation", phase)) %>%
  filter(str_detect(question, "rating")) %>%
  filter(str_detect(question, "baseline", negate = TRUE)) %>%
  mutate(answer = factor(answer, 
                         levels = 1:5, 
                         labels = likert_levels)) %>%
  pivot_wider(
    id_cols = c(phase, prolific_subject_id),
    names_from = question, 
    values_from = answer) %>% # keep only question columns %>%
  select(-c(phase, prolific_subject_id)) %>%
    mutate(across(everything(),
                ~ factor(.x, levels = likert_levels))) 

#plot
likert_plot_aware <- plot(likert(as.data.frame(likert_data_absolute_aware)), legend.position="right")

#just for ai blind 
likert_data_absolute_blind <- surveys_df %>%
  filter(condition == "blind") %>%
  select(phase, prolific_subject_id, question, answer) %>%
  mutate(phase = gsub("concept", "evaluation", phase)) %>%
  filter(str_detect(question, "rating")) %>%
  filter(str_detect(question, "baseline", negate = TRUE)) %>%
  mutate(answer = factor(answer, 
                         levels = 1:5, 
                         labels = likert_levels)) %>%
  pivot_wider(
    id_cols = c(phase, prolific_subject_id),
    names_from = question, 
    values_from = answer) %>% # keep only question columns %>%
  select(-c(phase, prolific_subject_id)) %>%
    mutate(across(everything(),
                ~ factor(.x, levels = likert_levels))) 

#plot

likert_plot_blind <- plot(likert(as.data.frame(likert_data_absolute_blind)), legend.position="right")

ggarrange(likert_plot_aware, likert_plot_blind, ncol = 1, labels = c("AI-Aware", "AI-Blind"))

```

```{r}
#| label: THESIS-RESULTS-FIGURE-ABSOLUTE-RATINGS-ALL-ITEMS
#| echo: false

combined_long_data <- surveys_df %>%
  mutate(phase = gsub("concept", "evaluation", phase)) %>%
  filter(str_detect(question, "rating")) %>%
  filter(str_detect(question, "baseline", negate = TRUE)) %>%
  # dlean names in long format
  mutate(question = str_to_title(str_replace_all(str_replace_all(question, "-rating", ""), "-", " "))) %>%
  mutate(question = ifelse(question == "Effectiveness", "Enjoyability", question)) %>% # i discuss effectiveness as enjoyability in the thesis 
  # set the desired order of questions (chose this order to demonstrate which dimensions were rated best)
  mutate(question = factor(question, levels = rev(c("Concept Creativity", "Concept Implementation", "Creativity", 
                                                    "Intentionality", "Authenticity", "Originality", "Enjoyability")))) %>%
  # handle condition factoring
  mutate(condition = factor(str_to_title(condition), levels = c("Aware", "Blind"), labels = c("AI-Aware", "AI-Blind"))) %>%
  # set response levels 
  mutate(answer = factor(answer, levels = 1:5, labels = likert_levels)) %>%
  # calculate percentages per condition and question
  group_by(question, condition, answer) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(question, condition) %>%
  mutate(percentage = count / sum(count))

abs_ratings_likert <- ggplot(combined_long_data, aes(x = percentage, y = question, fill = answer)) +
  
  geom_col(position = position_stack(reverse = TRUE), width = 0.7) +
  
  scale_fill_brewer(palette = "Blues", direction = 1) +
  scale_x_continuous(labels = scales::percent_format()) +
  facet_wrap(~ condition, ncol = 1) +
  labs(x = "Percentage of Responses", y = NULL, fill = "Response Rating") +
  theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    strip.background = element_blank(), 
    strip.text = element_text(color = "black", face = "bold", size = 14, hjust = 0),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    panel.grid.major.y = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.spacing = unit(2, "lines"), 
    legend.position = "bottom",
    legend.title = element_blank()
  ) +
  guides(fill = guide_legend(reverse = FALSE))

print(abs_ratings_likert)

```

# Linear mixed-models for creativity and related dimensions 
```{r}
#| label: big LMM for creativity analyses (RQ1 & RQ2)
#| echo: false

# function to run LMM and get marginal means for a specific dimension
run_creativity_lmm <- function(outcome) {
  # formula: outcome ~ condition + (1 | game_id) + (1 | prolific_subject_id)
  formula <- as.formula(paste(outcome, "~ condition + (1 | game_id) + (1 | prolific_subject_id)"))
  model <- lmer(formula, data = clean_analysis_df)
  
  # get marginal means
  em <- emmeans(model, ~ condition)
  
  return(list(model = model, emmeans = em))
}

# run models for all dimensions
outcomes <- c("creativity", "originality", "effectiveness", "authenticity", "intentionality")
results <- map(outcomes, run_creativity_lmm) %>% set_names(outcomes)

# extract marginal means for overall creativity (RQ1)
overall_creativity_means <- results$creativity$emmeans %>% as.data.frame()

#check statistical significance of overall creativity
contrast_creativity <- pairs(results$creativity$emmeans) %>% as.data.frame()

# repeat extracting marginal means for individual dimensions (RQ2)
individual_dim_means <- map_dfr(outcomes[-1], ~ {
  as.data.frame(results[[.x]]$emmeans) %>% mutate(dimension = .x)
})

contrast_individual_dims <- map_dfr(outcomes[-1], ~ {
  as.data.frame(pairs(results[[.x]]$emmeans)) %>% mutate(dimension = .x)
})

```

```{r}
#| label: individual LMMs to test where singularity is coming from 
#| echo: false 


creativity_model <- lmer(creativity ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)
originality_model <- lmer(originality ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)
effectiveness_model <- lmer(effectiveness ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)
authenticity_model <- lmer(authenticity ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)
intentionality_model <- lmer(intentionality ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)


creativity_emmeans <- emmeans(creativity_model, ~ condition)
originality_emmeans <- emmeans(originality_model, ~ condition)
effectiveness_emmeans <- emmeans(effectiveness_model, ~ condition)
authenticity_emmeans <- emmeans(authenticity_model, ~ condition)
intentionality_emmeans <- emmeans(intentionality_model, ~ condition)


#VarCorr(creativity_model)

#VarCorr(originality_model)

#VarCorr(effectiveness_model)

#VarCorr(authenticity_model)

#VarCorr(intentionality_model)
```

```{r}
#| label: calculating SD for each subject (investigating singularity)
#| include: false 

subject_sd_df <- clean_analysis_df %>%
  group_by(prolific_subject_id) %>%
  summarize(
    creativity_sd = sd(creativity, na.rm = TRUE),
    originality_sd = sd(originality, na.rm = TRUE),
    effectiveness_sd = sd(effectiveness, na.rm = TRUE),
    intentionality_sd = sd(intentionality, na.rm = TRUE),
    authenticity_sd = sd(authenticity, na.rm = TRUE),
    n_ratings = n()
  ) %>%
  ungroup()

ggplot(subject_sd_df, aes(x = creativity_sd)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Hist of Within-Participant SD of Creativity Ratings",
    x = "Participant Standard Deviation",
    y = "Number of Participants"
  )

ggplot(subject_sd_df, aes(x = originality_sd)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Hist of Within-Participant SD of Originality Ratings",
    x = "Participant Standard Deviation",
    y = "Number of Participants"
  )


ggplot(subject_sd_df, aes(x = effectiveness_sd)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Hist of Within-Participant SD of Effectiveness Ratings",
    x = "Participant Standard Deviation",
    y = "Number of Participants"
  )


ggplot(subject_sd_df, aes(x = intentionality_sd)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Hist of Within-Participant SD of Intentionality Ratings",
    x = "Participant Standard Deviation",
    y = "Number of Participants"
  )


ggplot(subject_sd_df, aes(x = authenticity_sd)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Hist of Within-Participant SD of Authenticity Ratings",
    x = "Participant Standard Deviation",
    y = "Number of Participants"
  )

```

```{r}
#| label: plotting stand devs across games for creativity and each creativity-relateed item (investigating singularity)
#| include: false 

game_sd_df <- clean_analysis_df %>%
  group_by(game_id) %>%
  summarize(
    creativity_sd = sd(creativity, na.rm = TRUE),
    originality_sd = sd(originality, na.rm = TRUE),
    effectiveness_sd = sd(effectiveness, na.rm = TRUE),
    intentionality_sd = sd(intentionality, na.rm = TRUE),
    authenticity_sd = sd(authenticity, na.rm = TRUE),
    n_ratings = n()
  ) %>%
  ungroup()

ggplot(game_sd_df, aes(x = creativity_sd)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Hist of Rating SDs Across Games (Creativity)",
    x = "SD of ratings per game",
    y = "Number of games"
  )

ggplot(game_sd_df, aes(x = originality_sd)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Hist of Rating SDs Across Games (Originality)",
    x = "SD of ratings per game",
    y = "Number of games"
  )

ggplot(game_sd_df, aes(x = effectiveness_sd)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Hist of Rating SDs Across Games (Effectiveness)",
    x = "SD of ratings per game",
    y = "Number of games"
  )

ggplot(game_sd_df, aes(x = intentionality_sd)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Hist of Rating SDs Across Games (Intentionality)",
    x = "SD of ratings per game",
    y = "Number of games"
  )

ggplot(game_sd_df, aes(x = authenticity_sd)) +
  geom_histogram(bins = 10) +
  labs(
    title = "Hist of Rating SDs Across Games (Authenticity)",
    x = "SD of ratings per game",
    y = "Number of games"
  )

```

```{r}
#| label: plotting dists of each participant's mean ratings of creativity and creativity-related items  (investigating singularity)
#| include: false 

participant_means <- clean_analysis_df %>%
  pivot_longer(
    cols = c(creativity, originality, effectiveness, authenticity, intentionality),
    names_to = "dimension",
    values_to = "rating"
  ) %>%
  group_by(prolific_subject_id, dimension) %>%
  summarize(mean_rating = mean(rating, na.rm = TRUE), .groups = "drop")

participant_sd_by_dimension <- participant_means %>%
  group_by(dimension) %>%
  summarize(sd_of_means = sd(mean_rating))

ggplot(participant_means, aes(x = mean_rating)) +
  geom_histogram(bins = 20) +
  facet_wrap(~dimension) +
  labs(
    title = "Distribution of Participant Mean Ratings by Dimension",
    x = "Participant Mean Rating",
    y = "Number of Participants"
  )


```

```{r}
#| label: boxplot of rating dists (bad for likert data; don't use)
#| include: false 

clean_analysis_df %>%
  pivot_longer(
    cols = c(creativity, originality, effectiveness, authenticity, intentionality),
    names_to = "dimension",
    values_to = "rating"
  ) %>%
  ggplot(aes(x = dimension, y = rating)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Ratings Across Dimensions",
    x = "Dimension",
    y = "Rating"
  )
```

```{r}
#| label: plotting each game's mean rating for each dimension  (investigating singularity)
#| include: false 

game_means <- clean_analysis_df %>%
  group_by(game_id) %>%
  summarize(
    creativity = mean(creativity),
    originality = mean(originality),
    effectiveness = mean(effectiveness),
    authenticity = mean(authenticity),
    intentionality = mean(intentionality)
  ) %>%
  pivot_longer(-game_id, names_to = "dimension", values_to = "mean_rating")

ggplot(game_means, aes(x = game_id, y = mean_rating)) +
  geom_point() +
  facet_wrap(~dimension) +
  coord_flip()
```

```{r}
#| label: old lmm report text 
#| echo: false

# 
# As seen in Figure 1, ratings of creativity spanned the whole Likert scale from 1 ("Not at all") to 5 ("Extremely creative") in both conditions. To get a better look at these ratings, I ran my pre-registered linear mixed-effects models predicting creativity ratings with condition (aware vs. blind) as a fixed effect and random intercepts for game and rater. This model had a singular fit with a participant-level random effect estimated variance of zero, indicating no detectable between-participant variability in ratings after accounting for game effects. Following standard practice, I deviated from my pre-registration by removing the participant random effect and retaining only a random intercept for game (rating \~ condition + (1 \| game_id)).
# 
# Marginal means estimated via the Kenward–Roger approximation indicated that creativity ratings fell around the mid-range of the creativity scale. These ratings were similar across the aware (*M* = `r round(overall_creativity_means$emmean[overall_creativity_means$condition == "aware"], 2)`, 95% CI \[`r round(overall_creativity_means$lower.CL[overall_creativity_means$condition == "aware"], 2)`, `r round(overall_creativity_means$upper.CL[overall_creativity_means$condition == "aware"], 2)`\]) and blind (*M* = `r round(overall_creativity_means$emmean[overall_creativity_means$condition == "blind"], 2)`, 95% CI \[`r round(overall_creativity_means$lower.CL[overall_creativity_means$condition == "blind"], 2)`, `r round(overall_creativity_means$upper.CL[overall_creativity_means$condition == "blind"], 2)`\]) conditions, and there was no statistically significant difference between ratings from AI-aware and AI-blind participants (*t*(`r round(contrast_creativity$df, 2)`) = `r round(contrast_creativity$t.ratio, 2)`, *p* = `r round(contrast_creativity$p.value, 3)`).

```


```{r}
#| label: tbl-absolute-creativity-dimension-results
#| include: false


results_table <- individual_dim_means %>%
  select(dimension, condition, emmean, lower.CL, upper.CL) %>%
  mutate(
    mean_ci = paste0(
      round(emmean, 2), " [",
      round(lower.CL, 2), ", ",
      round(upper.CL, 2), "]"
    )
  ) %>%
  select(dimension, condition, mean_ci) %>%
  pivot_wider(
    names_from = condition,
    values_from = mean_ci
  ) %>%
  left_join(
    contrast_individual_dims %>%
      mutate(
        t_stat = round(t.ratio, 2),
        df = round(df, 2),
        p = round(p.value, 3),
        test = paste0("t(", df, ") = ", t_stat)
      ) %>%
      select(dimension, test, p),
    by = "dimension"
  ) %>%
  rename(
    `Dimension` = dimension,
    `Aware M [95% CI]` = aware,
    `Blind M [95% CI]` = blind,
    `t-test` = test,
    `p` = p
  )

kable(results_table)
```

# PCA on creativity and related dimensions
```{r}
#| label: pca and biplot on dimensions 
#| include: false

# note that i haven't considered whether pca is a good fit here at all, was just curious to see the biplot
rating_matrix <- clean_analysis_df %>%
  select(creativity, originality, effectiveness, authenticity, intentionality) %>%
  rename(Creativity = creativity,
         Originality = originality,
         Enjoyability = effectiveness,
         Authenticity = authenticity,
         Intentionality = intentionality)

pca <- prcomp(rating_matrix, scale. = TRUE)

biplot(pca)

cumpve_endorsement <- summary(pca)[["importance"]]["Cumulative Proportion",]

plot(1:length(cumpve_endorsement), cumpve_endorsement, ylim = c(0,1), xlab = "PC", 
     ylab = "Cum Prop Variance Explained", type = "b", pch = 20)

summary(pca)
# 80% of variance accounted for by first principal component supports interrelatedness of these concepts
```

```{r}
#| label: biplot 
#| include: false

# fviz_pca_biplot(
#   pca,
#   repel = TRUE,            # avoids label overlap
#   col.var = "steelblue",   # variable arrows
#   col.ind = "gray70",      # points
#   alpha.ind = 0.5,
#   arrowsize = 0.8
# ) +
#   theme_minimal()


fviz_pca_var(
  pca,
  col.var = "steelblue",
  repel = TRUE
) +
  theme_minimal()


# fviz_pca_var(
#   pca,
#   col.var = "cos2",  # or "contrib"
#   gradient.cols = c("lightgray", "blue", "darkblue"),
#   repel = TRUE
# )


```

```{r}
#| label: THESIS-PCA-PLOT
#| echo: false

# make biplot from previous figure look nicer 
final_pca_plot <- fviz_pca_var(
  pca,
  col.var = "cos2", 
  gradient.cols = c("#A1B3D7", "#2A5699", "#08306B"), 
  
  repel = TRUE, 
  title = "",
  
  labelsize = 4.5,         
  box.padding = 0.8,       
  point.padding = 0.5    
) +
  theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    axis.title = element_text(face = "bold", size = 12, color = "black"),
    axis.text = element_text(size = 11, color = "black"),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.25),
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 11)
  ) +
labs(color = substitute(atop("Representation", "Quality (cos"^2*")")))
print(final_pca_plot)

```

# Spaghetti Plot
```{r}
#| label: fig-spaghetti-plots-participant-ratings
#| echo: false

# this plot is super cool but i don't really know if the complexity does enough to be worth it for readers

spaghetti_df <- clean_analysis_df %>%
  mutate(game_id = str_to_title(str_replace(game_id, "-[1-6]-", " ")))

ggplot(spaghetti_df,
       aes(x = game_id,
           y = creativity,
           group = prolific_subject_id)) +
  geom_line(alpha = .2) +
  labs(
    title = "Participant Rating Profiles Across Games",
    x = "Game",
    y = "Creativity Rating"
  ) +
    theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# plot with mean trend added
ggplot(spaghetti_df, aes(x = game_id, y = creativity)) +
  geom_line(aes(group = prolific_subject_id), alpha = 0.25, color = "black") +
  stat_summary(fun = mean, geom = "line", group = 1, color = "#2A5699", linewidth = 0.5) +
  labs(
    title = "Participant Rating Profiles (with Mean Trend)",
    x = "Game",
    y = "Creativity Rating"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# EDA correlations
```{r fig.width = 8, fig.height = 4}
#| label: EDA correlations between dimensions (scatterplot and cor matrix)
#| include: false

ggplot(clean_analysis_df, aes(x = creativity, y = originality)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition) 

ggplot(clean_analysis_df, aes(x = creativity, y = effectiveness)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = creativity, y = authenticity)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = creativity, y = intentionality)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = originality, y = effectiveness)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = originality, y = authenticity)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = originality, y = intentionality)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = effectiveness, y = authenticity)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = effectiveness, y = intentionality)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

ggplot(clean_analysis_df, aes(x = authenticity, y = intentionality)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap( ~condition)

rating_matrix <- clean_analysis_df %>%
  select(creativity, originality, effectiveness, authenticity, intentionality)

corrplot(cor(rating_matrix, use = "pairwise.complete.obs"),
         method = "color",
         type = "upper")
```

# Qual creativity 
```{r}
#| label: creativity qual analysis 
#| echo: false

# should i filter out free responses for games that were emergency exited? i don't think so. it probably also doesn't matter a whole lot. 

# read in data from coding it in google sheets 
creativity_evaluated_qual_df <- read.csv("coded_qual_data/creativity_evaluated_data.csv") 

# count how many times each theme was present for each tag and the total number of times
creativity_theme_counts <- creativity_evaluated_qual_df %>%
  pivot_longer(
    cols = c(
      novelty_originality,
      novelty_concept,
      novelty_mechanics,
      mechanics_gameplay,
      aesthetic_sensory_design,
      concept_theme,
      depth_complexity_meaning,
      simplicity_gameplay,
      clarity_learnability,
      engagement_fun,
      technical_execution,
      player_thinking_involvement_difficulty,
      depth_or_use_case_potential
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, tag) %>%
  mutate(n_pct = paste0(round(100*(n/243), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/243), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total))


# # see how many times some specific words that i flagged were present
# creativity_evaluated_qual_df %>%
# filter(str_detect(answer, regex("interesting", ignore_case = TRUE))) %>%
#   count()
# 
# creativity_evaluated_qual_df %>%
# filter(str_detect(answer, regex("instruction", ignore_case = TRUE))) %>%
#   count()
# 
# creativity_evaluated_qual_df %>%
# filter(str_detect(analyst_extra_note, regex("theme", ignore_case = TRUE))) %>%
#   count()
# 
# creativity_evaluated_qual_df %>%
# filter(str_detect(analyst_extra_note, regex("intentional", ignore_case = TRUE))) %>%
#   count()
# 
# creativity_evaluated_qual_df %>%
# filter(str_detect(answer, regex("random", ignore_case = TRUE))) %>%
#   count()
# 
# creativity_evaluated_qual_df %>%
# filter(str_detect(analyst_extra_note, regex("combination", ignore_case = TRUE))) %>%
#   count()



# count how many times each theme was present for each tag and the total number of times
surveys_df_unique <- surveys_df %>%
  distinct(prolific_subject_id, condition) %>%
  select(prolific_subject_id, condition)

creativity_theme_by_condition <- creativity_evaluated_qual_df %>%
  left_join(surveys_df_unique, by = "prolific_subject_id") %>%
  pivot_longer(
    cols = c(
      novelty_originality,
      novelty_concept,
      novelty_mechanics,
      mechanics_gameplay,
      aesthetic_sensory_design,
      concept_theme,
      depth_complexity_meaning,
      simplicity_gameplay,
      clarity_learnability,
      engagement_fun,
      technical_execution,
      player_thinking_involvement_difficulty,
      depth_or_use_case_potential
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, tag, condition) %>%
  mutate(n_pct = paste0(round(100*(n/121), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme, condition) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/121), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total))

wide_creativity_theme_by_condition <- creativity_theme_by_condition %>%
  select(theme, tag, condition, n, n_pct) %>%
  pivot_wider(
    names_from = condition,
    values_from = c(n, n_pct)
  )
```

```{r}
#| label: creativity qual on participant level
#| echo: false
#| warning: false
#| message: false 
# checking whether things were happening on a participant level

creativity_theme_counts_by_participant <- creativity_evaluated_qual_df %>%
  pivot_longer(
    cols = c(
      novelty_originality,
      novelty_concept,
      novelty_mechanics,
      mechanics_gameplay,
      aesthetic_sensory_design,
      concept_theme,
      depth_complexity_meaning,
      simplicity_gameplay,
      clarity_learnability,
      engagement_fun,
      technical_execution,
      player_thinking_involvement_difficulty,
      depth_or_use_case_potential
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, prolific_subject_id) %>%
  rename(n_tags_for_subject_theme_valence = n)

# see what percent of the time subjects had a theme appear in more than one of their free responses
creativity_theme_counts_by_participant_props <- creativity_theme_counts_by_participant %>%
count(n_tags_for_subject_theme_valence) %>%
  mutate(n_prop_sub_repeat_theme = paste0(round(100*(n / sum(n)), 0), "%"))

# if a good chunk of repeats, check which themes (curious if there are dominant themes)  
creativity_theme_counts_by_participant_repeats <- creativity_theme_counts_by_participant %>%
  filter(n_tags_for_subject_theme_valence > 1)  %>% # run code just to here to see whether it's predominantly 2s or 3s 
  count(theme) %>%
  rename(n_theme_tagged_two_three_times_per_subject = n) %>%
  mutate(prop_sample_repeated_theme = paste0(round(100*(n_theme_tagged_two_three_times_per_subject / sum(n_theme_tagged_two_three_times_per_subject)), 0), "%"))
```

# Baseline adjusted linear mixed-models
```{r}
#| label: baseline-adjusted creativity analyses
#| include: false

# function to run LMM with baseline covariate for a specific dimension
run_baseline_adjusted_lmm <- function(outcome, baseline_covariate) {
  # Formula: outcome ~ condition + baseline_covariate + (1 | game_id) + (1 | prolific_subject_id)
  formula <- as.formula(paste(outcome, "~ condition +", baseline_covariate, "+ (1 | game_id) + (1 | prolific_subject_id)"))
  model <- lmer(formula, data = clean_analysis_df)
  
  # Get marginal means for condition, averaged over the baseline covariate
  em <- emmeans(model, ~ condition)
  
  return(list(model = model, emmeans = em))
}

# Define mapping between dimensions and their corresponding baseline covariates
dim_baseline_pairs <- list(
  originality = "baseline_originality_rating",
  effectiveness = "baseline_effectiveness_rating",
  authenticity = "baseline_authenticity_rating",
  intentionality = "baseline_intentionality_rating"
)

# run baseline-adjusted models
baseline_adjusted_results <- imap(dim_baseline_pairs, ~ run_baseline_adjusted_lmm(.y, .x))

# extract results for all dimensions 
baseline_adjusted_means <- imap_dfr(baseline_adjusted_results, ~ {
  as.data.frame(.x$emmeans) %>% mutate(dimension = .y)
})

# extract significance for all dimensions 
contrast_individual_baseline_adjusted_dims <- map_dfr(outcomes[-1], ~ {
  as.data.frame(pairs(baseline_adjusted_results[[.x]]$emmeans)) %>% mutate(dimension = .x)
})

```

```{r}
#| label: tbl-baseline-adjusted 
#| include: false

baseline_results_table <- baseline_adjusted_means %>%
  select(dimension, condition, emmean, lower.CL, upper.CL) %>%
  mutate(
    mean_ci = paste0(
      round(emmean, 2), " [",
      round(lower.CL, 2), ", ",
      round(upper.CL, 2), "]"
    )
  ) %>%
  select(dimension, condition, mean_ci) %>%
  pivot_wider(
    names_from = condition,
    values_from = mean_ci
  ) %>%
  left_join(
    contrast_individual_baseline_adjusted_dims %>%
      mutate(
        t_stat = round(t.ratio, 2),
        df = round(df, 2),
        p = round(p.value, 3),
        test = paste0("t(", df, ") = ", t_stat)
      ) %>%
      select(dimension, test, p),
    by = "dimension"
  ) %>%
  rename(
    Dimension = dimension,
    `Aware M [95% CI]` = aware,
    `Blind M [95% CI]` = blind,
    `t-test` = test,
    p = p
  )

kable(baseline_results_table)
```

```{r}
#| label: old baseline results report text
#| echo: false

# To account for potential differences in participants’ baseline evaluations, we fit an additional linear mixed-effects model predicting creativity ratings with condition (aware vs. blind) as a fixed effect, baseline rating as a covariate, and random intercepts for game and rater. Baseline-adjusted marginal means estimated via the Kenward–Roger approximation indicated that creativity ratings remained similar across the aware (*M* = `r round(baseline_adjusted_means$emmean[baseline_adjusted_means$condition=="aware"],2)`, 95% CI \[`r round(baseline_adjusted_means$lower.CL[baseline_adjusted_means$condition=="aware"],2)`, `r round(baseline_adjusted_means$upper.CL[baseline_adjusted_means$condition=="aware"],2)`\]) and blind (*M* = `r round(baseline_adjusted_means$emmean[baseline_adjusted_means$condition=="blind"],2)`, 95% CI \[`r round(baseline_adjusted_means$lower.CL[baseline_adjusted_means$condition=="blind"],2)`, `r round(baseline_adjusted_means$upper.CL[baseline_adjusted_means$condition=="blind"],2)`\]) conditions, and this difference was not statistically significant, *t*(`r round(contrast_individual_baseline_adjusted_dims$df,2)`) = `r round(contrast_individual_baseline_adjusted_dims$t.ratio,2)`, *p* = `r round(contrast_individual_baseline_adjusted_dims$p.value,3)`.


# As seen in Figure 1, ratings for specific dimensions of creativity also spanned the entire scale, where 1 was "Not at all" and 5 was "Extremely \[dimension\]". I fit my pre-registered linear mixed-effects model for originality, effectiveness, authenticity, and intentionality ratings, with condition (aware vs. blind) as a fixed effect and random intercepts for game and rater. Marginal means estimated via the Kenward–Roger approximation indicated that ratings all generally fell within the mid-range of the Likert scale for each dimension, and were similar across the aware and blind conditions for all four dimensions (@tbl-absolute-creativity-dimension-results). None of the aware–blind contrasts were statistically significant (all *ps* \> .05).

```

# Table all LMM results
```{r}
#| label: tbl-creativity-dimensions-absolute-and-adjusted
#| include: false

# combining both tables into one for easy display

# add column to tables saying whether adjusted or unadjusted
stacked_results_table <- bind_rows(
  results_table %>%
    mutate(Model = "Observed") %>%
    select(
      Dimension,
      Model,
      `Aware M [95% CI]`,
      `Blind M [95% CI]`,
      `t-test`,
      p
    ),
  
  baseline_results_table %>%
    mutate(Model = "Adjusted") %>%
    select(
      Dimension,
      Model,
      `Aware M [95% CI]`,
      `Blind M [95% CI]`,
      `t-test`,
      p
    )
) # %>%
 # arrange(Dimension, Model)

kable(stacked_results_table)
```

```{r}
#| label: THESIS-FULL-LLM-RESULTS-TABLE
#| echo: false 

# nicer looking version of the previous table
thesis_table_data <- stacked_results_table %>%
  mutate(Dimension = str_to_title(Dimension)) %>%
  mutate(Dimension = factor(Dimension, levels = c("Originality", "Effectiveness", "Authenticity", "Intentionality"))) %>%
  arrange(Dimension, desc(Model)) %>%
  mutate(p = case_when(
    as.numeric(p) < 0.001 ~ "< .001",
    TRUE ~ sub("^0", "", sprintf("%.3f", as.numeric(p)))
  )) %>%
  select(
    Dimension,
    Model,
    `M [95% CI]` = `Aware M [95% CI]`,
    `M [95% CI] ` = `Blind M [95% CI]`, 
    `t-test`,
    `p` = p
  )

kable(thesis_table_data, 
      format = "html", 
      align = c("l", "l", "c", "c", "c", "c"),
      booktabs = TRUE,
      linesep = "") %>% 
  
  kable_classic(full_width = FALSE, 
                position = "center",
                html_font = "serif") %>% 
  
  add_header_above(c(" " = 2, "AI-Aware" = 1, "AI-Blind" = 1, " " = 2)) %>%
  collapse_rows(columns = 1, valign = "top") %>%
  
  row_spec(0, extra_css = "padding-top: 8px; padding-bottom: 8px;") %>%
    row_spec(c(1, 3, 5, 7), 
           extra_css = "padding-top: 8px; padding-bottom: 8px; border-bottom: 1px solid #CCCCCC;") %>%
  
  row_spec(c(2, 4, 6), 
           extra_css = "padding-top: 8px; padding-bottom: 8px; border-bottom: 1px solid black;") %>%
  
    row_spec(8, extra_css = "padding-top: 8px; padding-bottom: 8px;")
```

# Likert baseline
```{r}
#| label: fig-likert-plots-baseline-ratings
#| include: false
#| message: false
#| warning: false

baseline_df_wide <- baseline_df %>%
  pivot_longer(
    cols = starts_with("baseline_"), # select columns to pivot
    names_to = "Dimension",          # name of the new column for category names
    values_to = "Rating"             # name of the new column for ratings
  ) %>%
  mutate("Dimension" = as.character(Dimension),
         Rating = as.numeric(Rating))


# likert plot! 
# put df into likert format 
likert_data_baseline <- baseline_df %>%
  select( # putting in defined order and removing prolific_sub_id
    baseline_originality_rating,
    baseline_effectiveness_rating,
    baseline_intentionality_rating,
    baseline_authenticity_rating
  ) %>%
  mutate(across(everything(),
                ~ factor(.x, levels = 1:5, labels = likert_levels)))

#plot
likert_plot_baseline <- plot(likert(as.data.frame(likert_data_baseline)), legend.position="right")

#just for ai aware 
likert_data_baseline_aware <- baseline_df %>%
  filter(prolific_subject_id %in% aware_ids) %>%
  select( # putting in defined order and removing prolific_sub_id
    baseline_originality_rating,
    baseline_effectiveness_rating,
    baseline_intentionality_rating,
    baseline_authenticity_rating
  ) %>%
  mutate(across(everything(),
                ~ factor(.x, levels = 1:5, labels = likert_levels)))

#plot
likert_plot_baseline_aware <- plot(likert(as.data.frame(likert_data_baseline_aware)), legend.position="right", ordered = FALSE)

likert_plot_baseline_aware$items <- factor(
  likert_plot_baseline_aware$items,
  levels = c(
    "baseline_originality_rating",
    "baseline_effectiveness_rating",
    "baseline_intentionality_rating",
    "baseline_authenticity_rating"
  )
)

#just for ai blind 
likert_data_baseline_blind <- baseline_df %>%
  filter(prolific_subject_id %in% blind_ids) %>%
  select( # putting in defined order and removing prolific_sub_id
    baseline_originality_rating,
    baseline_effectiveness_rating,
    baseline_intentionality_rating,
    baseline_authenticity_rating
  ) %>%
  mutate(across(everything(),
                ~ factor(.x, levels = 1:5, labels = likert_levels)))

#plot
likert_plot_baseline_blind <- plot(likert(as.data.frame(likert_data_baseline_blind)), legend.position="right", ordered = FALSE)

likert_plot_baseline_blind$items <- factor(
  likert_plot_baseline_blind$items,
  levels = c(
    "baseline_originality_rating",
    "baseline_effectiveness_rating",
    "baseline_intentionality_rating",
    "baseline_authenticity_rating"
  )
)

ggarrange(likert_plot_baseline_aware, likert_plot_baseline_blind, ncol = 1, labels = c("AI-Aware", "AI-Blind"))

```

```{r}
#| label: THESIS-RESULTS-FIGURE-BASELINE
#| echo: false

likert_levels <- c("Not at all", "Somewhat", "Moderately", "Very", "Extremely")
baseline_desired_order <- rev(c("Intentionality", "Authenticity", "Originality", "Enjoyability"))

combined_baseline_long <- baseline_df %>%
  mutate(condition = case_when(
    prolific_subject_id %in% aware_ids ~ "AI-Aware",
    prolific_subject_id %in% blind_ids ~ "AI-Blind",
    TRUE ~ NA_character_ # Just in case there are IDs not in either list
  )) %>%
  select(condition, 
         baseline_intentionality_rating, 
         baseline_authenticity_rating, 
         baseline_originality_rating, 
         baseline_effectiveness_rating) %>%
  pivot_longer(
    cols = -condition,
    names_to = "question",
    values_to = "answer"
  ) %>%
  mutate(question = str_replace_all(question, "baseline_", "")) %>%
  mutate(question = str_replace_all(question, "_rating", "")) %>%
  mutate(question = str_to_title(question)) %>%
  mutate(question = ifelse(question == "Effectiveness", "Enjoyability", question)) %>%  # Apply explicit ordering
  mutate(question = factor(question, levels = baseline_desired_order)) %>%
  mutate(condition = factor(condition, levels = c("AI-Aware", "AI-Blind"))) %>%
  mutate(answer = factor(answer, levels = 1:5, labels = likert_levels)) %>%
  group_by(question, condition, answer) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(question, condition) %>%
  mutate(percentage = count / sum(count))

baseline_final_plot <- ggplot(combined_baseline_long, aes(x = percentage, y = question, fill = answer)) +
  
  geom_col(position = position_stack(reverse = TRUE), width = 0.7) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  scale_x_continuous(labels = scales::percent_format()) +
  facet_wrap(~ condition, ncol = 1) +
  labs(x = "Percentage of Responses", y = NULL, fill = "Response Rating") +
  theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    strip.background = element_blank(), 
    strip.text = element_text(color = "black", face = "bold", size = 14, hjust = 0),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    panel.grid.major.y = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.spacing = unit(2, "lines"), 
    legend.position = "bottom",
    legend.title = element_blank()
  ) +
  guides(fill = guide_legend(reverse = FALSE))

print(baseline_final_plot)

```

```{r fig.width = 6, fig.height = 7}
#| label: another version of likert plot
#| include: false

baseline_long <- baseline_df %>%
  mutate(condition = case_when(
    prolific_subject_id %in% aware_ids ~ "aware",
    prolific_subject_id %in% blind_ids ~ "blind"
  )) %>%
  select(condition, starts_with("baseline_")) %>%
  pivot_longer(
    cols = starts_with("baseline_"),
    names_to = "dimension",
    values_to = "response"
  )

baseline_long$response <- factor(
  baseline_long$response,
  levels = 1:5,
  labels = likert_levels
)

baseline_summary <- baseline_long %>%
  group_by(condition, dimension, response) %>%
  summarize(n = n(), .groups = "drop") %>%
  group_by(condition, dimension) %>%
  mutate(pct = n / sum(n))

ggplot(baseline_summary,
       aes(x = dimension, y = pct, fill = response)) +
  geom_bar(stat = "identity", width = 0.7) +
  coord_flip() +
  facet_wrap(~condition, ncol = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  labs(
    x = NULL,
    y = "Percent of responses",
    fill = "Rating"
  ) +
#  theme_minimal(base_size = 13) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold")
  )

```

```{r}
#| label: fig-likert-differences-absolute-baseline-ungrouped
#| include: false 

likert_difference_levels <- c("-5 (LLM Games < Baseline)", "-4", "-3", "-2", "-1", "0 (LLM Games = Baseline)", "1", "2", "3", "4", "5 (LLM Games > Baseline)")

likert_difference_aware <- difference_ratings_df %>%
  filter(condition == "aware") %>%
  select(starts_with("difference")) %>%
  mutate(across(everything(),
                ~ factor(.x, levels = -5:5, labels = likert_difference_levels)))


likert_plot_difference_aware <- plot(likert(as.data.frame(likert_difference_aware)), legend.position="right", ordered = FALSE)


likert_difference_blind <- difference_ratings_df %>%
  filter(condition == "blind") %>%
  select(starts_with("difference")) %>%
  mutate(across(everything(),
                ~ factor(.x, levels = -5:5, labels = likert_difference_levels)))

likert_plot_difference_blind <- plot(likert(as.data.frame(likert_difference_blind)), legend.position="right", ordered = FALSE)

ggarrange(likert_plot_difference_aware, likert_plot_difference_blind, ncol = 1)

```

```{r}
#| label: fig-likert-differences-absolute-baseline-grouped
#| echo: false 

likert_difference_levels_categorical <- c("-5 (LLM Games < Baseline)", "-4", "-3", "-2", "-1", "0 (LLM Games = Baseline)", "1", "2", "3", "4", "5 (LLM Games > Baseline)")

likert_difference_grouped <- difference_ratings_df %>%
  select(condition, prolific_subject_id, starts_with("difference")) %>%
  group_by(prolific_subject_id) %>%
  mutate(mean_difference_originality = mean(difference_originality),
         mean_difference_effectiveness = mean(difference_effectiveness),
         mean_difference_authenticity = mean(difference_authenticity),
         mean_difference_intentionality = mean(difference_intentionality)
         ) %>%
  ungroup(prolific_subject_id) %>%
  distinct(prolific_subject_id, .keep_all = TRUE) %>%
  select(condition, starts_with("mean_difference")) %>%
  pivot_longer(
    cols = starts_with("mean_difference"),
    names_to  = "rating_item",
    values_to = "rating"
 )

likert_difference_ggplot <- ggplot(likert_difference_grouped, aes(x = rating_item, y = rating)) +
  geom_boxplot() +
  facet_wrap(~condition, ncol = 1) +
  labs(title = "Dists of diff in LLM game rating vs general games rating", subtitle = "Means were grouped by subject so interpretations can be on the subject level", x = "Creativity dimension", y = "LLM Game Rating - Baseline Rating (-5 to 5)") +
  coord_flip() +
  geom_hline(yintercept = 0, color = "blue") 

likert_difference_ggplot

```

# Qual originality
```{r}
#| label: originality qual analysis
#| echo: false 
#| warning: false
#| message: false

originality_evaluated_qual_data <- read.csv("coded_qual_data/originality_evaluated_data.csv")

# main qual df 
originality_theme_counts <- originality_evaluated_qual_data %>%
  pivot_longer(
    cols = c(
      gameplay_mechanics,
      visual_aesthetics,
      theme_narrative_concept,
      audio_sound_music,
      comparative_reference,
      surprise_unexpectedness,
      player_thinking_challenge,
      improve_technical_execution,
      too_confusing_buggy_execution_to_fully_evaluate
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  # group_by(prolific_subject_id) %>% # can use this and distinct to see different by-subject levels of the counts
 # distinct(prolific_subject_id, .keep_all = TRUE) %>%
  filter(tag > 0) %>%
  count(theme, tag) %>%
  mutate(n_pct = paste0(round(100*(n/243), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/243), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total)) 



# look at diffs btwn conditions
originality_theme_by_condition <- originality_evaluated_qual_data %>%
  left_join(surveys_df_unique, by = "prolific_subject_id") %>%
  pivot_longer(
    cols = c(
      gameplay_mechanics,
      visual_aesthetics,
      theme_narrative_concept,
      audio_sound_music,
      comparative_reference,
      surprise_unexpectedness,
      player_thinking_challenge,
      improve_technical_execution,
      too_confusing_buggy_execution_to_fully_evaluate
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, tag, condition) %>%
  mutate(n_pct = paste0(round(100*(n/121), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme, condition) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/121), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total))

wide_originality_theme_by_condition <- originality_theme_by_condition %>%
  select(theme, tag, condition, n, n_pct) %>%
  pivot_wider(
    names_from = condition,
    values_from = c(n, n_pct)
  )
```

```{r}
#| label: originality qual on participant level
#| echo: false
#| warning: false
#| message: false 
# checking whether things were happening on a participant level

originality_theme_counts_by_participant <- originality_evaluated_qual_data %>%
  pivot_longer(
    cols = c(
      gameplay_mechanics,
      visual_aesthetics,
      theme_narrative_concept,
      audio_sound_music,
      comparative_reference,
      surprise_unexpectedness,
      player_thinking_challenge,
      improve_technical_execution,
      too_confusing_buggy_execution_to_fully_evaluate
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, prolific_subject_id) %>%
  rename(n_tags_for_subject_theme_valence = n)

# see what percent of the time subjects had a theme appear in more than one of their free responses
originality_theme_counts_by_participant_ <- originality_theme_counts_by_participant %>%
count(n_tags_for_subject_theme_valence) %>%
  mutate(n_prop_sub_repeat_theme = paste0(round(100*(n / sum(n)), 0), "%"))

# if a good chunk of repeats, check which themes (curious if there are dominant themes)  
originality_theme_counts_by_participant_repeats <- originality_theme_counts_by_participant %>%
  filter(n_tags_for_subject_theme_valence > 1)  %>% # run code just to here to see whether it's predominantly 2s or 3s 
  count(theme) %>%
  rename(n_theme_tagged_two_three_times_per_subject = n) %>%
  mutate(prop_sample_repeated_theme = paste0(round(100*(n_theme_tagged_two_three_times_per_subject / sum(n_theme_tagged_two_three_times_per_subject)), 0), "%"))
```

# Qual enjoyability (effectiveness)

```{r}
#| label: effectiveness qual analysis
#| echo: false 

effectiveness_evaluated_qual_data <- read.csv("coded_qual_data/effectiveness_evaluated_data.csv")

effectiveness_theme_counts <- effectiveness_evaluated_qual_data %>%
  pivot_longer(
    cols = c(
      mechanics,
      sub_mechanics_execution,
      sub_mechanics_design,
      visuals_aesthetic,
      audio,
      general_no_specification,
      difficult,
      simple,
      confusion,
      bugs,
      concept,
      broken_bugs_couldnt_evaluate_at_all
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  # group_by(prolific_subject_id) %>% # can use this and distinct to see different by-subject levels of the counts
 # distinct(prolific_subject_id, .keep_all = TRUE) %>%
  filter(tag > 0) %>%
  count(theme, tag) %>%
  mutate(n_pct = paste0(round(100*(n/243), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/243), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total)) 


# look at diffs btwn conditions
effectiveness_theme_by_condition <- effectiveness_evaluated_qual_data %>%
  left_join(surveys_df_unique, by = "prolific_subject_id") %>%
  pivot_longer(
    cols = c(
      mechanics,
      sub_mechanics_execution,
      sub_mechanics_design,
      visuals_aesthetic,
      audio,
      general_no_specification,
      difficult,
      simple,
      confusion,
      bugs,
      concept,
      broken_bugs_couldnt_evaluate_at_all
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, tag, condition) %>%
  mutate(n_pct = paste0(round(100*(n/121), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme, condition) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/121), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total))

wide_effectiveness_theme_by_condition <- effectiveness_theme_by_condition %>%
  select(theme, tag, condition, n, n_pct) %>%
  pivot_wider(
    names_from = condition,
    values_from = c(n, n_pct)
  )

```

```{r}
#| label: effectiveness qual on participant level
#| echo: false
#| warning: false
#| message: false 
# checking whether things were happening on a participant level

effectiveness_by_participant <- effectiveness_evaluated_qual_data %>%
  pivot_longer(
    cols = c(
      mechanics,
      sub_mechanics_execution,
      sub_mechanics_design,
      visuals_aesthetic,
      audio,
      general_no_specification,
      difficult,
      simple,
      confusion,
      bugs,
      concept,
      broken_bugs_couldnt_evaluate_at_all
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, prolific_subject_id) %>% # count number of times each subject tagged each theme positively (tag = 1) and each theme negatively (tag = 2)
  rename(n_tags_for_subject_theme_valence = n)

# see what percent of rows had a theme appear in more than one of their free responses
effectiveness_by_participant_summary <- effectiveness_by_participant %>%
count(n_tags_for_subject_theme_valence) %>%
  mutate(n_prop_sub_repeat_theme = paste0(round(100*(n / sum(n)), 0), "%")) 

# if a good chunk of repeats, check which themes (curious if there are dominant themes)  
effectiveness_theme_counts_by_participant_repeats <- effectiveness_by_participant %>%
  filter(n_tags_for_subject_theme_valence > 1)  %>% # run code just to here to see whether it's predominantly 2s or 3s 
  count(theme) %>% # for each theme, count how many participants mentioned this theme more than once 
  rename(n_participants_tagged_theme_two_three_times = n) %>%
  mutate(prop_sample_repeated_theme = paste0(round(100*(n_participants_tagged_theme_two_three_times / total_num_clean_subjects_in_data), 0), "%"))

```

# Qual authenticity and intentionality

```{r}
#| label: authenticity intentionality qual analysis
#| echo: false 

authint_evaluated_qual_df <- read.csv("coded_qual_data/authenticity_intentionality_evaluated_data.csv")


authint_theme_counts <- authint_evaluated_qual_df %>%
  pivot_longer(
    cols = c(
      originality,
      overall_idea_concept,
      fun_engagement_challenging,
      visual,
      mechanics_gameplay,
      completeness_execution,
      design_intent,
      personal_invest_effort,
      accessibility,
      ai_gen
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  # group_by(prolific_subject_id) %>% # can use this and distinct to see different by-subject levels of the counts
 # distinct(prolific_subject_id, .keep_all = TRUE) %>%
  filter(tag > 0) %>%
  count(theme, tag) %>%
  mutate(n_pct = paste0(round(100*(n/243), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/243), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total)) 



# look at diffs btwn conditions
originality_theme_by_condition <- originality_evaluated_qual_data %>%
  left_join(surveys_df_unique, by = "prolific_subject_id") %>%
  pivot_longer(
    cols = c(
      gameplay_mechanics,
      visual_aesthetics,
      theme_narrative_concept,
      audio_sound_music,
      comparative_reference,
      surprise_unexpectedness,
      player_thinking_challenge,
      improve_technical_execution,
      too_confusing_buggy_execution_to_fully_evaluate
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  filter(tag > 0) %>%
  count(theme, tag, condition) %>%
  mutate(n_pct = paste0(round(100*(n/121), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme, condition) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/121), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total))

wide_originality_theme_by_condition <- originality_theme_by_condition %>%
  select(theme, tag, condition, n, n_pct) %>%
  pivot_wider(
    names_from = condition,
    values_from = c(n, n_pct)
  )

```

```{r}
#| label: ai aware qual analysis
#| echo: false

ai_aware_evaluated_qual_data <- read.csv("coded_qual_data/ai_aware_evaluated_data.csv")

ai_aware_theme_counts <- ai_aware_evaluated_qual_data %>%
  mutate(valence = case_when(
    valence == "positively_affected" ~ 1,
    valence == "negatively_affected" ~ 2,
    valence == "mixed" ~ 3,
    valence == "no_effect" ~ 4,
    valence == "unclear" ~ 5
      )) %>%
  pivot_longer(
    cols = c(
      valence,
      lowered_expectations_standards,
      respect_impressed_AI_capability,
      AI_as_legitimate_tool,
      devaluation_due_to_AI_origin,
      quality_deficits_linked_to_AI,
      human_oversight_needed,
      wasnt_aware_of_ai_generation,
      judged_game_merit,
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  # group_by(prolific_subject_id) %>% # can use this and distinct to see different by-subject levels of the counts
 # distinct(prolific_subject_id, .keep_all = TRUE) %>%
  filter(tag > 0) %>%
  count(theme, tag) %>%
  mutate(n_pct = paste0(round(100*(n/40), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/40), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total)) 

```

```{r}
#| label: ai blind qual analysis
#| echo: false


ai_blind_evaluated_qual_data <- read.csv("coded_qual_data/ai_blind_evaluated_data.csv")

ai_blind_theme_counts <- ai_blind_evaluated_qual_data %>%
  mutate(valence = case_when(
    valence == "positive" ~ 1,
    valence == "negative" ~ 2,
    valence == "mixed" ~ 3,
    valence == "no_effect" ~ 4,
    valence == "unclear" ~ 5
      )) %>%
  pivot_longer(
    cols = c(
      valence,
      prior_AI_detection_suspicion,
      judged_game_merit,
      diminished_authenticity,
      reduced_effort_intentionality,
      AI_explains_quality_issues,
      AI_creativity,
      positive_AI_regard
    ),
    names_to = "theme",
    values_to = "tag"
  ) %>%
  # group_by(prolific_subject_id) %>% # can use this and distinct to see different by-subject levels of the counts
 # distinct(prolific_subject_id, .keep_all = TRUE) %>%
  filter(tag > 0) %>%
  count(theme, tag) %>%
  mutate(n_pct = paste0(round(100*(n/41), 0), "%")) %>% # turn into percent for easy reporting 
  group_by(theme) %>%
  mutate(n_total = sum(n), # grab total sum
         n_total_pct = paste0(round(100*(n_total/41), 0), "%")) %>%# turn into percent for easy reporting
  arrange(desc(n_total)) 

```

# Game concept

```{r}
#| label: game concept analyses (RQ3)
#| echo: false

# analyze concept creativity and implementation accuracy using the LMM framework
concept_creativity_lmm <- lmer(concept_creativity ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)
implementation_accuracy_lmm <- lmer(implementation_accuracy ~ condition + (1 | game_id) + (1 | prolific_subject_id), data = clean_analysis_df)

# marginal means for concept creativity and implementation accuracy
concept_creativity_means <- emmeans(concept_creativity_lmm, ~ condition) %>% as.data.frame()
implementation_accuracy_means <- emmeans(implementation_accuracy_lmm, ~ condition) %>% as.data.frame()

# comparison: Concept Creativity vs. Game Creativity (within-participant)
# pivot long for Rating Type (Concept vs. Game)
long_comparison_df <- clean_analysis_df %>%
  pivot_longer(
    cols = c(creativity, concept_creativity),
    names_to = "rating_type",
    values_to = "creativity_rating"
  ) %>%
  mutate(rating_type = factor(rating_type, levels = c("concept_creativity", "creativity")))

# repeated-measures LMM with within-participant factor of Rating Type
rating_type_comparison_lmm <- lmer(
  creativity_rating ~ rating_type * condition + (1 | game_id) + (1 | prolific_subject_id),
  data = long_comparison_df
)

# interaction: Rating Type x Condition
rating_type_means <- emmeans(rating_type_comparison_lmm, ~ rating_type * condition) %>% as.data.frame()

# exploratory moderation: Does implementation accuracy affect the relationship between concept creativity and game creativity?
moderation_lmm <- lmer(
  creativity ~ concept_creativity * implementation_accuracy + (1 | game_id) + (1 | prolific_subject_id),
  data = clean_analysis_df
)


# look at results 
#summary(concept_creativity_lmm)
#summary(implementation_accuracy_lmm)
#concept_creativity_means
#implementation_accuracy_means

#summary(rating_type_comparison_lmm)
#rating_type_means

# main effect of rating type
#emmeans(rating_type_comparison_lmm, pairwise ~ rating_type)

# interaction breakdown
#emmeans(rating_type_comparison_lmm, pairwise ~ rating_type | condition)

# look at whether implementation moderates concept to real game ratings 
#summary(moderation_lmm)

# probe the interaction: simple slopes of concept_creativity 
# at low, medium, and high implementation accuracy
#emtrends(moderation_lmm, ~ implementation_accuracy, 
 #        var = "concept_creativity",
  #       at = list(implementation_accuracy = c(1, 3, 5)))
```
